
Inspired by Model Predictive Path Integral (MPPI)~\citep{williams2015model} control, we perform iterative latent-space trajectory optimization during inference.
Instead of fixing latent priors to standard Gaussians, we adaptively update the noise distributions of the video and value modules based on trajectory evaluations.
We maintain two independent Gaussian noise distributions:
a video flow noise distribution
$
\scriptsize
f_{\mathrm{vid}}^{(k)} = \mathcal{N}\!\left(
\boldsymbol{\mu}_{\mathrm{vid}}^{(k)},
\operatorname{diag}\!\left((\boldsymbol{\sigma}_{\mathrm{vid}}^{(k)})^2\right)
\right),
$
and a trajectory value flow noise distribution
$
\scriptsize
f_{\mathrm{val}}^{(k)} = \mathcal{N}\!\left(
\boldsymbol{\mu}_{\mathrm{val}}^{(k)},
\operatorname{diag}\!\left((\boldsymbol{\sigma}_{\mathrm{val}}^{(k)})^2\right)
\right),
$
where $k \in \{1,\dots,K\}$ denotes the iteration index.

At iteration $k$, we first sample $M$ groups of video latent noise:
$
\scriptstyle
\left\{
\mathbf{z}_{\mathrm{vid}}^{(m)}
\right\}_{m=1}^M
\sim
f_{\mathrm{vid}}^{(k-1)}.
$
Each latent sample is denoised by the video generation module $\mathcal{W}$ to produce video feature chunking:

$
\scriptstyle
\mathbf{x}^{(m)}_{1:H_1} = \mathcal{W}\!\big(\mathbf{z}_{\mathrm{vid}}^{(m)}\big)$,
where $\mathbf{x}^{(m)} \in \mathbb{R}^{H_1 \times d_v}$ represents a predicted future visual trajectory feature.
For each video feature chunking $\mathbf{x}^{(m)}$, we sample $N$ groups of trajectory value flow noise:
$
\scriptstyle
\left\{
\mathbf{z}_{\mathrm{val}}^{(m,n)}
\right\}_{n=1}^N
\sim
f_{\mathrm{val}}^{(k-1)},
$
and denoise them using the trajectory value module $\mathcal{V}$ conditioned on the same video features:

$
\scriptstyle
\mathbf{v}^{(m,n)} = \mathcal{V}\!\big(\mathbf{x}^{(m)}, \mathbf{z}_{\mathrm{val}}^{(m,n)}\big)$,
where $\mathbf{v}^{(m,n)}$ denotes a predicted state-value.

To obtain a stable assessment of future trajectory quality, we compute the Signal-to-Noise Ratio (SNR) of the value predictions for each exploration:

$\scriptstyle \operatorname{SNR}^{(m,n)} = \mathbb{E}\!\left[\mathbf{v}^{(m,n)}\right] / \left(\operatorname{Std}\!\left[\mathbf{v}^{(m,n)}\right] + \epsilon\right)$
where $\epsilon$ is a small constant for numerical stability.
The final score of the $m^{th}$ video exploration is defined as the most reliable evaluation among its $N$ value estimates:

$
\scriptstyle
\phi^{(m)} = \max_{n \in \{1,\dots,N\}} \operatorname{SNR}^{(m,n)}$.
Based on the scores $\{\phi^{(m)}\}_{m=1}^M$, we select the top $K_1$ elite video explorations.
The video noise distribution is updated using the empirical mean and standard deviation of their corresponding latent variables:

{
\scriptsize
\begin{equation}
\begin{aligned}
\boldsymbol{\mu}_{\mathrm{vid}}^{(k)}
&=
\frac{1}{K_1}
\sum_{m \in \mathcal{E}_{\mathrm{vid}}}
\mathbf{z}_{\mathrm{vid}}^{(m)}, \text{and }
\boldsymbol{\sigma}_{\mathrm{vid}}^{(k)}
=
\sqrt{
\frac{1}{K_1}
\sum_{m \in \mathcal{E}_{\mathrm{vid}}}
\left(
\mathbf{z}_{\mathrm{vid}}^{(m)} -
\boldsymbol{\mu}_{\mathrm{vid}}^{(k)}
\right)^2
}.
\end{aligned}
\label{eq:video_mu_sigma}
\end{equation}
}
Similarly, from all $M \times N$ value evaluations, we select the top $K_2$ samples with the highest SNR and update the value noise distribution:

{\scriptsize
\begin{equation}
\begin{aligned}
\boldsymbol{\mu}_{\mathrm{val}}^{(k)}
&=
\frac{1}{K_2}
\sum_{(m,n) \in \mathcal{E}_{\mathrm{val}}}
\mathbf{z}_{\mathrm{val}}^{(m,n)},
 \text{and }
\boldsymbol{\sigma}_{\mathrm{val}}^{(k)}
=
\sqrt{
\frac{1}{K_2}
\sum_{(m,n) \in \mathcal{E}_{\mathrm{val}}}
\left(
\mathbf{z}_{\mathrm{val}}^{(m,n)} -
\boldsymbol{\mu}_{\mathrm{val}}^{(k)}
\right)^2
}.
\end{aligned}
\label{eq:value_mu_sigma}
\end{equation}
}

To prevent distribution collapse and improve iteration stability, we apply variance decay and exponential smoothing:

\(
\scriptstyle
\boldsymbol{\mu}^{(k)} \!\leftarrow\!
\alpha \boldsymbol{\mu}^{(k)} + (1-\alpha)\boldsymbol{\mu}^{(k-1)},\;
\boldsymbol{\sigma}^{(k)} \!\leftarrow\!
\beta \boldsymbol{\sigma}^{(k)} + (1-\beta)\boldsymbol{\sigma}^{(k-1)}
\)
with smoothing parameters $\alpha, \beta \in (0,1)$.
After $K$ iterations, we sample latent variables from the optimized distributions
$f_{\mathrm{vid}}^{(K)}$ and $f_{\mathrm{val}}^{(K)}$,
and obtain the corresponding video and value features via flow denoising.
These features are finally fused by the action decoding module to predict decision-making actions.
Algo.~\ref{alg:latent_mppi} summarizes the proposed latent planning procedure.

